---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figs/README-"
)
```

#Radiomics: Texture Analysis Matrices

The `radiomics` package is a set of tools for computing texture matrices and features from images. 

The release version of this package is available from [CRAN](https://cran.r-project.org/web/packages/radiomics/index.html) using:

```{r,eval=FALSE}
install.packages("radiomics")
```

Or you can install the development version of the package using:

```{r,echo=FALSE, message=FALSE}
devtools::load_all(".")
```
```{r, eval=FALSE}
devtools::install_github("joelcarlson/radiomics")

library(radiomics)
```

#Texture Matrices

In the package are functions for calculating four different types of matrices used to quantify the texture of an image. 

These matrices are the:

  - Gray Level Co-occurrence Matrix
  - Gray Level Run Length Matrix
  - Gray Level Size Zone Matrix
  - Multiple Gray Level Size Zone Matrix

Detailed descriptions and references for each matrix can be found [here](http://joelcarlson.me/2015/07/10/radiomics-package/)
Detailed usage directions for calculating features and matrices can be found in the package vignette (use `browseVignettes(package = "radiomics")`)

#Using the Package

##Building Texture Matrices

Texture matrices can be created from 2D images by using the abbreviated matrix name as a function call:

```{r, eval=FALSE}
tumor <- radiomics::tumor #2D MRI slice of a brain tumor
glcm(tumor)
glrlm(tumor)
glszm(tumor)
mglszm(tumor)
```

A matrix with the class of the texture matrix type is return, as shown here using `glcm(tumor, n_grey=4)`

```{r,echo=FALSE}
glcm(tumor, n_grey=4)
```
```{r}
class(glcm(tumor, n_grey=4))[1]
```


##Visualizing Texture Matrices

Each matrix type has an associated `image` function for visualization of the results:

```{r,eval=FALSE}
image(glcm(tumor))
image(glrlm(tumor))
image(glszm(tumor))
image(mglszm(tumor))
```

The `image` functions make use of the `viridis` scale, as shown here using `image(glcm(tumor, n_grey=64))`:

```{r tumorglcm,echo=FALSE, message=FALSE}
image(glcm(tumor, n_grey=64))
```

##Calculating Features

Each matrix type has an associated `calc_features` function, which returns an object of class `data.frame` with a single observation for each calculated feature. First order features can also be calculated on 2D matrices.

```{r,eval=FALSE}
calc_features(tumor)
calc_features(glcm(tumor))
calc_features(glrlm(tumor))
calc_features(glszm(tumor))
calc_features(mglszm(tumor))
```



